---
title: "Final Model: Analyzing Movie Revenue "
author: "Andrew Batmunkh, Kevin Cheung"
date: "`r Sys.Date()`"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

library(tidyverse)
library(dplyr)
library(knitr)
```

## Data Source

TMDB 5000 Movie Dataset (Kaggle)  
URL: https://www.kaggle.com/datasets/tmdb/tmdb-movie-metadata

## 1. Import and Cleaning
**"movies" is the original preliminary model with no transformation, where "movies2" is the log10 transformed model.**

```{r import, message=FALSE}
movies <- read_csv("tmdb_5000_movies.csv") %>%
  mutate(
    release_year  = as.integer(substr(release_date, 1, 4)),
    primary_genre = stringr::str_match(genres, '"name"\\s*:\\s*"([^"]+)"')[,2],
    is_english    = as.integer(original_language == "en"),       
    has_homepage  = as.integer(!is.na(homepage) & homepage != "")
  ) %>%
  filter(budget > 0, revenue > 0, !is.na(primary_genre))

cat("Cleaned dataset contains", nr`row(movies), "observations and", 
    ncol(movies), "variables.\n")

movies2 <- read_csv("tmdb_5000_movies.csv") %>%
  mutate(
    release_year   = as.integer(substr(release_date, 1, 4)),
    primary_genre  = stringr::str_match(genres, '"name"\\s*:\\s*"([^"]+)"')[,2],
    
    log_budget     = ifelse(budget > 0, log10(budget), NA_real_),
    log_revenue    = ifelse(revenue > 0, log10(revenue), NA_real_),
    log_vote_count = ifelse(vote_count > 0, log10(vote_count), NA_real_),
    
    is_english     = as.integer(original_language == "en"),
    has_homepage   = as.integer(!is.na(homepage) & homepage != "")
  ) %>%
  filter(budget > 0, revenue > 0, vote_count > 0, !is.na(primary_genre))

cat("Cleaned dataset contains", nrow(movies2), "observations and", 
    ncol(movies2), "variables.\n")

```

## 2a. Numerical Summary for prelim model
```{r summary}
num_vars <- c("budget","runtime","popularity","vote_average", "vote_count","release_year")

num_summary <- movies %>%
  select(all_of(num_vars)) %>%
  pivot_longer(everything(), names_to = "variable", values_to = "x") %>%
  group_by(variable) %>%
  summarise(
    n      = sum(!is.na(x)),
    mean   = mean(x, na.rm = TRUE),
    sd     = sd(x, na.rm = TRUE),
    min    = min(x, na.rm = TRUE),
    q1     = quantile(x, .25, na.rm = TRUE),
    median = median(x, na.rm = TRUE),
    q3     = quantile(x, .75, na.rm = TRUE),
    max    = max(x, na.rm = TRUE),
    outliers = {
      q1_ <- quantile(x, .25, na.rm = TRUE); q3_ <- quantile(x, .75, na.rm = TRUE)
      iqr_ <- q3_ - q1_; sum(x < (q1_ - 1.5*iqr_) | x > (q3_ + 1.5*iqr_), na.rm = TRUE)
    }
  ) %>%
  mutate(across(where(is.double), ~ round(.x, 2)))

print.data.frame(num_summary, row.names = FALSE)
```
## 2b. Numerical Summary for final transformed model
```{r}
num_vars <- c(
  "budget","runtime","popularity","vote_average","vote_count","release_year",
  "log_budget","log_revenue","log_vote_count"
)

num_summary2 <- movies2 %>%
  select(all_of(num_vars)) %>%
  pivot_longer(everything(), names_to = "variable", values_to = "x") %>%
  group_by(variable) %>%
  summarise(
    n      = sum(!is.na(x)),
    mean   = mean(x, na.rm = TRUE),
    sd     = sd(x, na.rm = TRUE),
    min    = min(x, na.rm = TRUE),
    q1     = quantile(x, .25, na.rm = TRUE),
    median = median(x, na.rm = TRUE),
    q3     = quantile(x, .75, na.rm = TRUE),
    max    = max(x, na.rm = TRUE),
    outliers = {
      q1_ <- quantile(x, .25, na.rm = TRUE); 
      q3_ <- quantile(x, .75, na.rm = TRUE)
      iqr_ <- q3_ - q1_
      sum(x < (q1_ - 1.5*iqr_) | x > (q3_ + 1.5*iqr_), na.rm = TRUE)
    }
  ) %>%
  mutate(across(where(is.double), ~ round(.x, 2)))

print.data.frame(num_summary2, row.names = FALSE)
```

```{r categorical-summary}
cat("\nPrimary Genre Distribution:\n")  
table(movies$primary_genre)
cat("\n\nEnglish Language Movies:\n") 
table(movies$is_english)
cat("\n\nMovies with Homepage:\n")  
table(movies$has_homepage)
```
##3. ## 3. Final transformation Multiple Linear Regression Model
```{r}
fit_log <- lm(log_revenue ~ log_budget + runtime + popularity + vote_average + log_vote_count + release_year + is_english + has_homepage + primary_genre + budget:primary_genre, data = movies2)

summary(fit_log)
```

## 3. Preliminary Multiple Linear Regression Model

```{r model-fit}
fit_revenue <- lm(revenue ~ budget + runtime + popularity + vote_average + vote_count + release_year + is_english + has_homepage + primary_genre + budget:primary_genre, data = movies)

summary(fit_revenue)
```

## 4a. Residual Analysis for final transformation model

```{r}
par(mfrow = c(1,2))
plot(fit_log)
```


## 4b. Residual Analysis for the preliminary model

```{r residual-plots, fig.height=8, fig.width=8}
par(mfrow = c(2,2))
plot(fit_revenue)
```


## AIC/BIC analysis
```{r}
# AIC and BIC for both models
AIC(fit_revenue)
BIC(fit_revenue)

AIC(fit_log)
BIC(fit_log)
```

